#  Working with relational databases

```{r}
library(palmerpenguins)
library(tidyverse)

penguins %>% 
  tidyr::pivot_longer(contains("_"), # measurement cols
                      names_to = c("variable_name"), 
                      values_to = "value")
```

```{r}
penguins_long <- penguins %>% 
  tidyr::pivot_longer(contains("_"), # measurement cols
                      names_to = c("part", "measure", "unit"), 
                      names_sep = "_")

penguins_long
```


```{r}
penguins_long %>% 
  tidyr::pivot_wider(names_from = c("part", "measure", "unit"), # pivot these columns
                     values_from = "value", # take the values from here
                     names_sep = "_") # combine col names using an underscore
```


```{r}
# give each measured penguin a number
penguins_with_id <- penguins %>%  mutate(id=seq(nrow(.)))

# make the data long again
penguins_long_ided <- penguins_with_id %>% 
      tidyr::pivot_longer(contains("_"), # measurement cols
                      names_to = c("part", "measure", "unit"), 
                      names_sep = "_")

# and wide again
penguins_long_ided2 <-  penguins_long_ided %>% 
    tidyr::pivot_wider(names_from = c("part", "measure", "unit"), 
                     values_from = "value", 
                     names_sep = "_") 
```





```{r}
penguins %>% head()
```


```{r}
penguins %>% mutate(penguin_id = row_number(),.before = species)
```


```{r}
food_preference <-
  tibble(
    penguin_species = c("Adelie", "Chinstrap", "Gentoo"),
    favorite_food = c("peanut butter", "cheese", "chocolate")
  )
```

```{r}
left_join(
  penguins, food_preference, by = c("species" = "penguin_species")
)
```

```{r}
penguins %>% mutate(penguin_id = row_number(),.before = species)
```



Let’s refresh your memory. Take the penguins dataset.

As we all know, Adelie penguins love peanut butter, Chinstrap penguins prefer cheese and Gentoo penguins do anything for chocolate.

Here is a table describing food preference:




```{r}
food_preference <-
  tibble(
    penguin_species = c("Adelie", "Chinstrap", "Gentoo"),
    favorite_food = c("peanut butter", "cheese", "chocolate")
  )
```


```{r}
joined <-  left_join(
  penguins, food_preference, by = c("species" = "penguin_species")
)
```


Now make a new tibble containing only the information on penguin species, sex and bill length measured in 2009 using filter and select.

#penguins_long_ided
#species island    sex     year    id bill_length_mm bill_depth_mm


```{r}

##filter(year=="2009")
# species island    sex     year    id bill_length_mm bill_depth_mm


head(penguins_long_ided2)  

new_tbl <- penguins_long_ided2  %>%  filter(year==2009) %>%  select(species,sex,bill_length_mm)

head(new_tbl)
```

